---
title: "Low Insulin Diet Study Report"
author: "Alex"
date: "April 24, 2025"
output: html_document
---

<style>
h2 {
  font-size: 28px;
  font-weight: bold;
}
h3 {
  font-size: 20px;
  font-weight: bold;
}
pre, code {
  font-size: 14px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(broom)
```

Introduction

This report analyzes a 12-week diet study with 50 participants, comparing four diet types (Keto, Low-Carb, Standard, Control) and the effect of Ozempic on weight loss, HOMA-IR, and fasting insulin. The dataset was cleaned using Google BigQuery and analyzed in R.

Data Loading

```{r load-data}

data <- read_csv("low_insulin_diet_bigquery_cleaned.csv")
head(data)
```

Weight Loss Analysis

```{r weight-analysis}

weight_change <- data %>%
  group_by(participant_id, diet_type, ozempic_use) %>%
  summarise(
    weight_start = weight_kg[week == 0],
    weight_end = weight_kg[week == 12],
    weight_loss = weight_start - weight_end
  ) %>%
  ungroup()

weight_summary <- weight_change %>%
  group_by(diet_type, ozempic_use) %>%
  summarise(
    avg_weight_loss = mean(weight_loss, na.rm = TRUE),
    sd_weight_loss = sd(weight_loss, na.rm = TRUE),
    n = n()
  ) %>%
  ungroup()

weight_summary
```

Weight Loss Visualization

```{r weight-plot, fig.width=8, fig.height=6}

ggplot(weight_summary, aes(x = diet_type, y = avg_weight_loss, fill = ozempic_use)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Average Weight Loss by Diet Type and Ozempic Use",
       x = "Diet Type",
       y = "Average Weight Loss (kg)") +
  theme_minimal()
```

HOMA-IR and Fasting Insulin Analysis

```{r homa-ir-analysis}

homa_ir_change <- data %>%
  group_by(participant_id, diet_type, ozempic_use) %>%
  summarise(
    homa_ir_start = homa_ir[week == 0],
    homa_ir_end = homa_ir[week == 12],
    homa_ir_change = homa_ir_start - homa_ir_end
  ) %>%
  ungroup()

homa_ir_summary <- homa_ir_change %>%
  group_by(diet_type, ozempic_use) %>%
  summarise(
    avg_homa_ir_change = mean(homa_ir_change, na.rm = TRUE),
    sd_homa_ir_change = sd(homa_ir_change, na.rm = TRUE),
    n = n()
  ) %>%
  ungroup()

homa_ir_summary
```

Conclusion

Keto and Low-Carb diets resulted in greater weight loss and reductions in HOMA-IR and fasting insulin compared to Standard and Control diets. Ozempic use enhanced these effects, particularly for Keto participants.
Additional visualizations are available on Tableau Public: https://public.tableau.com/app/profile/alexander.gallagher/viz/LowInsulinDietStudyVisualizations/DietStudyVisualizations.









